Project Report\\\textsc{DEVS Ex Machina}

### 2015-05-24 manual/projectrapport.tex

\section*{Terminology}
The literature sometimes exchanges terms, to avoid any confusion the following frequent terms are listed with their associated meaning as used in this document:
\begin{itemize}
  \item Core : alias kernel. Single instance, containing a set of models it is responsible for, exchanges messages generated by those models and keeps time state. Can correspond with Logical Process in terms of what it represents.
  \item Logical Process : set of models with strong internal coherence.
  \item Distributed parallelism : computation distributed over several physically separated machines communicating by means of a network stack.
  \item SMP : symmetric multiprocessing, running a parallel computation on several distinct CPU cores/threads on the same machine with shared memory (and/or cache).
  \item message : any exchange of high level information between entities, user defined or for inter-kernel synchronization.
  \item lookahead : Timespan relative to current state where the output of the model is not influenced by its input. (as per definition in \cite{cons}, app B)
  \item eot : Earliest output time. Timepoint in future where the core/kernel will output events.
  \item eit : Earliest input time. Timepoint in future where the core/kernel expects to
  be change state on receiving event.
\end{itemize}
Lookahead is collected per model per kernel, eit/eot are based on information from models but determined on a core/kernel level.

\chapter{Planning}
\section{Parallel}
\begin {itemize}
\item GVT integration : week 0-1, Ben and Tim
\item Conservative : week 2-3 , Ben and Tim
\end {itemize}
\section{General}
\begin {itemize}
\item Dynamic Structured : week 1-3 Stijn
\item DevStone and examples : week 0-2 Matthijs and Pieter
\item Serialization : week 0 Pieter
\item Simulation control : week 1-3 Matthijs and Pieter  
\item Tracers : week 0 Stijn
\item Benchmarking PyPDevs Adevs : week 3 Matthijs and Pieter
\end {itemize}
\section{Optimization}
Week 4 , general review.

@@

\section{Operating system support}
The project requires a compliant C++11 compiler and library, but beyond that is not really bound to special features. For testing, GTest is a requirement, and the only two other dependencies are the ubiquitous Boost and Cereal, the last is provided with the project as it is header only.

\section{Object representation}
In Python everything is a reference, and since the DEVS formalisms require that the user be allowed to override most of the basic entities, the best representation for an object was a shared\textunderscore ptr. Passing by value could have been safer (e.g. between threads) but would introduce slicing.\\ Another advantage was that we would use as little memory as possible, for example sending a message means passing an const reference to a shared pointer, which is usually equal in size to 2 pointers. By passing by const ref we minimize most of the penalty usually involved in copyconstructing shared pointers. \\
The only downside is the synchronized access to the reference count in the shared pointer. There is also a need for more boilerplate code to use shared pointers in containers (providing pass-through hash function, comparison etc).

@@

\section{Serialization}
For serialization we used the Cereal \cite{cereal} header-only framework. We have used separate load and save functions for full flexibility. For each type that occurs in the form of a pointer we implemented a static load and construct function since the object first has to be created before the saved data can be imported. Cereal has a small coupling with the code (models) that has to be written by the user as the user has to register his own classes due to polymorphism.

>>>>

\section*{Terminology}
The literature sometimes exchanges terms, to avoid any confusion the following frequent terms are listed with their associated meaning as used in this document:
\begin{itemize}
  \item Core : alias kernel. Single instance, containing a set of models for which it is responsible, exchanges messages generated by those models and keeps time state. Can correspond with Logical Process in terms of what it represents.
  \item Logical Process : set of models with strong internal coherence.
  \item Distributed parallelism : computation distributed over several physically separated machines communicating by means of a network stack.
  \item SMP : symmetric multiprocessing, running a parallel computation on several distinct CPU cores/threads on the same machine with shared memory (and/or cache).
  \item message : any exchange of high level information between entities, user defined or for inter-kernel synchronization.
  \item lookahead : Timespan relative to current state where the output of the model is not influenced by its input. (as per definition in \cite{cons}, app B)
  \item eot : Earliest output time. Timepoint in future where the core/kernel will output events.
  \item eit : Earliest input time. Timepoint in future where the core/kernel expects to
  be change state on receiving event.
\end{itemize}
Lookahead is collected per model per kernel, eit/eot are based on information from models but determined on a core/kernel level.

\chapter{Planning}
\section{Parallel}
\begin {itemize}
\item GVT integration : week 0-1, Ben and Tim
\item Conservative : week 2-3 , Ben and Tim
\end {itemize}
\section{General}
\begin {itemize}
\item Dynamic Structured : week 1-3 Stijn
\item DevStone and examples : week 0-2 Matthijs and Pieter
\item Serialization : week 0 Pieter
\item Simulation control : week 1-3 Matthijs and Pieter  
\item Tracers : week 0 Stijn
\item Benchmarking PyPDevs Adevs : week 3 Matthijs and Pieter
\end {itemize}
\section{Optimization}
Week 4 general review.

@@

\section{Operating system support}
The project requires a compliant C++11 compiler and library, but beyond that is not really bound to special features. For testing, GTest is a requirement, and the only two other dependencies are the ubiquitous Boost and Cereal, the latter is provided with the project as it is header only.

\section{Object representation}
In Python everything is a reference, and since the DEVS formalisms require that the user be allowed to override most of the basic entities, the best representation for an object was a shared\textunderscore ptr. Passing by value could have been safer (e.g. between threads) but would introduce slicing.\\ Another advantage was that we would use as little memory as possible, for example sending a message means passing a const reference to a shared pointer, which is usually equal in size to 2 pointers. By passing by const ref we minimize most of the penalty usually involved in copyconstructing shared pointers. \\
The only downside is the synchronized access to the reference count in the shared pointer. There is also a need for more boilerplate code to use shared pointers in containers (providing pass-through hash function, comparison etc).

@@

\section{Serialization}
For serialization we used the Cereal \cite{cereal} header-only framework. We opted for separate load and save functions for full flexibility. For each type that occurs in the form of a pointer we implemented a static load and construct method since the object first has to be created before the saved data can be imported. Cereal has a small coupling with the code (models) that has to be written by the user as the user has to register his own classes due to polymorphism.

All the main functionality to load objects is implemented in the serialize function itself. Most classes also have an implementation for load and construct so that objects can separately be constructed before filling it with the saved data. This is necessary in cases where pointers to objects of those classes have to be serialized or where there is no default constructor available. The implementation of those load and construct methods is always restricted to creating an object with - most of the time - meaningless temporary datamembers after which the serialize function of the created object is called to load the actual data.

Since Cygwin is not compatible with the std::to\_string() function, which is used for serialization to human readable formats like JSON or XML, only serialization to binary files is supported. Even in the binary archive implementation of Cereal the std::to\_string() function occurred, though only on non-decisive places, resulting in some slightly modified headers for this project.  The binary archives are lighter and give more speed than the user-readable formats but comes with the disadvantage of not providing full garanty that saved data can be correctly loaded on another architecture.

Since Cereal was hard to use and gave a lot of undesired errors while serializing const datamembers and objects of purely virtual classes, we decided to violate some proper design rules to overcome this.

As we have experienced problems with an unressolved bug from Cereal - covering the errors in case of specific nested circular smart pointers - we were forced to not always follow the elegant design of Cereal and explicit force the calling of some serialization function in the load and construct functions.

### 2015-05-24 manual/tex/reqs.tex

\section{System requirements}
\subsubsection{Compiler}
The project requires a recent version of a standards-compliant C++11 compiler and runtime. The following are compilers/runtimes used to develop the simulator (and therefore guaranteed to work)
\begin{itemize}
  \item G++ 4.8.2 + libstdc++
  \item G++ 4.9.2 + libstdc++
  \item Clang 3.4.0 + libstdc++
  \item Clang 3.5.0 + libstdc++
  \item G++ 4.9.2 on Cygwin $>=$ 1.7
\end{itemize}
Especially the C++11 threading support is essential.
\subsubsection{Dependencies}
The build system requires CMake $>=$ 2.8, the presence of standard build tools is assumed (make etc.).
The following are the functional dependencies:
\begin{itemize}
  \item Cereal (source included)
  \item GTest $>=$ 1.6
  \item Boost heap $>=$ 1.55  
\end{itemize}
Boost can be installed on *nix with the following command:\\
Debian based:\\
\texttt{\$sudo apt-get install libboost-all-dev}\\
Red-hat based:\\
\texttt{\$sudo yum install boost-devel}\\
\subsubsection{Preparing the build}
We use thread-sanitizer to verify thread-safety at runtime instrumented by the compiler (g++ and clang). This however requires building the project (and by extension any linked libraries) with position independent flags. For the simulator this is resolved by CMake, however for gtest you'll need to manually build GTest with this extra flag:\\
In file internal\textunderscore utils.cmake change compile options to append the compile flag \texttt{-fpie} :
\texttt{set(cxx\textunderscore base\textunderscore flags "-Wall -Wshadow -fpie")}\\
If you have any non-standard locations where the compiler should search for Boost and/or GTest, place them in the file localpreferences.txt.\\
\texttt{SET(\{GTEST|BOOST\}\_ROOT "your path here")}. Make sure that for gtest, the listed path contains both the "include" folder and libraries in a dedicated folder "lib". For Boost the path needs to contain the "include" folder with the boost headers.
\paragraph{NOTE} If you use Mac OS X, the project links with the \texttt{libc++} library by default. GTest does not. To prevent linker errors when compiling the project you should build the GTest libraries with the flag \texttt{-stdlib=libc++}.
\subsubsection{Building}
On *nix, a build script is provided to drive CMake:\\
\texttt{\$main/generate\_build.sh \{Debug|Release\} \{g++|clang++\} }\\
Alternatively, invoke cmake manually : \\
\texttt{\$mkdir build }\\
\texttt{\$cd build}\\
\texttt{\$cmake -G"Eclipse CDT4 - Unix Makefiles" -DCMAKE\_CXX\_COMPILER\_ARG1=\\"\{-std=c++11|-std=gnu++11\}" -DCMAKE\_CXX\_COMPILER="\{clang++|g++\}" \\-DCMAKE\_BUILD\_TYPE=\{Release|Debug\} ../main}\\
\texttt{\$make -j\{CPU-CORES\}}\\

On Windows the default script should work but if you come into any problems, the commands given above work as well.
\subsubsection{Development}
The cmake invocation has created an Eclipse CDT project for you to open the simulator code in:
\begin{itemize}
  \item Open Eclipse Luna (with CDT plugin)
  \item File, Import, General
  \item Select "Existing projects into Workspace"
  \item navigate to the root folder of the project (containing ./main, ./build)
\end{itemize}

>>>>

\section{System requirements}
\subsubsection{Compiler}
The project requires a recent version of a standards-compliant C++11 compiler and runtime. The following are compilers/runtimes used to develop the simulator (and therefore guaranteed to work)
\begin{itemize}
  \item G++ 4.8.2 + libstdc++
  \item G++ 4.9.2 + libstdc++
  \item Clang 3.4.0 + libstdc++
  \item Clang 3.5.0 + libstdc++
  \item G++ 4.9.2 on Cygwin $>=$ 1.7
\end{itemize}
Especially the C++11 threading support is essential.
\subsubsection{Dependencies}
The build system requires CMake $>=$ 2.8, the presence of standard build tools is assumed (make etc.).
The following are the functional dependencies:
\begin{itemize}
  \item Cereal (source included)
  \item GTest $>=$ 1.6
  \item Boost heap $>=$ 1.55  
\end{itemize}
Boost can be installed on *nix with the following command:\\
Debian based:\\
\texttt{\$sudo apt-get install libboost-all-dev}\\
Red-hat based:\\
\texttt{\$sudo yum install boost-devel}\\
\subsubsection{Preparing the build}
We use thread-sanitizer to verify thread-safety at runtime instrumented by the compiler (g++ and clang). This however requires building the project (and by extension any linked libraries) with position-independent flags. For the simulator this is resolved by CMake, however for gtest you'll need to manually build GTest with this extra flag:\\
In file internal\textunderscore utils.cmake change compile options to append the compile flag \texttt{-fpie} :
\texttt{set(cxx\textunderscore base\textunderscore flags "-Wall -Wshadow -fpie")}\\
If you have any non-standard locations where the compiler should search for Boost and/or GTest, place them in the file localpreferences.txt.\\
\texttt{SET(\{GTEST|BOOST\}\_ROOT "your path here")}. Make sure that for gtest, the listed path contains both the "include" folder and libraries in a dedicated folder "lib". For Boost the path needs to contain the "include" folder with the boost headers.
\paragraph{NOTE} If you use Mac OS X, the project links with the \texttt{libc++} library by default. GTest does not. To prevent linker errors when compiling the project you should build the GTest libraries with the flag \texttt{-stdlib=libc++}.
\subsubsection{Building}
On *nix, a build script is provided to drive CMake:\\
\texttt{\$main/generate\_build.sh \{Debug|Release\} \{g++|clang++\} }\\
Alternatively, invoke cmake manually : \\
\texttt{\$mkdir build }\\
\texttt{\$cd build}\\
\texttt{\$cmake -G"Eclipse CDT4 - Unix Makefiles" -DCMAKE\_CXX\_COMPILER\_ARG1=\\"\{-std=c++11|-std=gnu++11\}" -DCMAKE\_CXX\_COMPILER="\{clang++|g++\}" \\-DCMAKE\_BUILD\_TYPE=\{Release|Debug\} ../main}\\
\texttt{\$make -j\{CPU-CORES\}}\\

On Windows the default script should work but if you come into any problems, the commands given above work if you use the CXX compiler argument \texttt{-std=gnu++11}.
\subsubsection{Development}
The cmake invocation has created an Eclipse CDT project for you to open the simulator code in:
\begin{itemize}
  \item Open Eclipse Luna (with CDT plugin)
  \item File, Import, General
  \item Select "Existing projects into Workspace"
  \item navigate to the root folder of the project (containing ./main, ./build)
\end{itemize}
